Rapid evolution of a conspicuous dynamic visual signal in an introduced chameleon

### 2021-05-31 ms.Rmd
Transferred over Martin's revisions, fixed up figures

---
title: "Evolution of a conspicuous dynamic visual signal in an introduced chameleon"
author: Whiting et al. 
date: "`r Sys.Date()`"
bibliography: refs.bib
csl: science.csl
output: 
  bookdown::word_document2:
    toc: no
    toc_depth: 6
    number_sections: false
    reference_docx: template.docx
  bookdown::html_document2:
    code_folding: hide
    number_sections: no
    toc: yes
    toc_depth: 6
    toc_float: yes
    bibliography: refs.bib
editor_options: 
  chunk_output_type: console
---

@@

```{r getlibrary, echo = FALSE, results = "hide"}
# load libraries
devtools::install_github("karthik/wesanderson")
pacman::p_load(lme4, lmerTest, ggplot2, lsmeans, emmeans, multcomp, dplyr, wesanderson, patchwork, latex2exp, png, cowplot, magick, jpeg, clubSandwich, flextable)
source("./R/func.R")

@@

    # Check robustness of results to non-indepenedence using Robust Variance Estimator (RVE)
    Display_1_IDs <- Display_1 %>% filter(!is.na(sc_SVL), !is.na(Population)) # filter out misssing data
       robust_ConspicDisplayAC <- coef_test(ConspicDisplayAC, 
                                            cluster = Display_1_IDs$ID, vcov = "CR2")  
    robust_ConspicDisplayAC_CI <- conf_int(ConspicDisplayAC, 
                                           cluster = Display_1_IDs$ID, vcov = "CR2")

@@

 # Check robustness of results to non-indepenedence using Robust Variance Estimator (RVE)
    Courtship_IDs <- Courtship %>% filter(!is.na(sc_SVL), !is.na(Population)) # filter out misssing data
    robust_ConspicCourtshipAC <- coef_test(ConspicCourtshipAC, 
                                           cluster = Courtship_IDs$ID, vcov = "CR2")
 robust_ConspicCourtshipAC_CI <- conf_int(ConspicCourtshipAC, 
                                          cluster = Courtship_IDs$ID, vcov = "CR2")

@@

  # Check robustness of results to non-indepenedence using Robust Variance Estimator (RVE)
  robust_localAdapt_DispBgrdAC <- clubSandwich::conf_int(DispBgrdAC, cluster = DisplayBgrd$ID, vcov =  "CR2")

@@

Martin J. Whiting $^{1,*}$, Brenden S. Holland $^{2}$, J. Scott Keogh$^{3}$, Daniel W.A.Noble$^{3}$, Katrina J. Rankin$^{4}$ and Devi M. Stuart-Fox$^{4}$

$^{1}$ Department of Biological Sciences, Macquarie University, Sydney, NSW 2109, Australia

# Abstract
Rapid evolution frequently follows a significant change to an organism’s environment, such as during colonization of a novel habitat. Vertebrates, including Caribbean Anolis lizards and Galapagos finches, have demonstrated a remarkable capacity for morphological change of functional traits (limb length, beak size) on ecological time scales. Traits that are sexually selected can be similarly affected, although well-documented examples from wild populations are rare. The Jackson’s chameleon (*Chamaeleo x. jacksonii*) from East Africa was inadvertently introduced to the Hawaiian island of Oahu in 1972. In its native range, Jackson’s chameleons are preyed upon by a wide range of predators, including snakes and birds, which are thought to shape the dynamics of color change in chameleons. We took advantage of the Hawaiian introduction, an environment with only a few potential predators and therefore, very low predation pressure, to test for character release of conspicuous social color signals. To test this hypothesis, we measured color change in response to multiple social and predator stimuli in wild chameleons in both Hawai’i and Kenya. We used visual modelling of colour signals to estimate signal conspicuousness to both the chameleon and predator (bird, snake) visual systems. The Hawai’i and Kenya backgrounds were a similar color, but the major difference was in brightness: the Hawai’i background was darker. As a result, chameleons from both Kenya and Hawai’i were brighter against the Hawai’i background. However, Hawai’i males were more conspicuous (courtship, display) against their own background for some body regions (mainly gular), providing evidence for character release. Also, Hawai’i males were more conspicuous (courtship, display) to snake and bird predators against their own background. We suggest that relaxing natural selection can result in signal release in systems experiencing strong sexual selection.

@@

When Darwin tackled the joint problems of speciation and evolutionary change through natural selection, he was of the view that evolution was a gradual process that took place over extended time periods. This view understandably persisted for at least another 120 years following the publication of On the Origin of Species (Darwin, 1859). We now have numerous examples of evolutionary change following natural selection in both the laboratory and wild, and not only in invertebrates with short generation times. Vertebrates, including Caribbean Anolis lizards and Galapagos finches, have demonstrated a remarkable capacity for morphological change of functional traits (limb length, beak size) on ecological time scales (e.g. @losos97). Furthermore, a far more recent development since Darwin’s seminal work on natural selection is the recognition that sexual selection can also drive significant changes in phenotype over short time intervals.

Conspicuous colour signals evolve because of the fitness advantage conferred to the signaller, and frequently becomes increasingly elaborate through a combination of male contest competition and female preference (e.g. runaway selection). However, trait exaggeration is typically halted by natural selection because there is a threshold beyond which conspicuousness is too costly. Evolution can respond to this selection pressure in a variety of ways that include shifting conspicuous signals to areas of the body that might be concealed to predators. For example, many conspicuous lizards have colourful throats and venters that they flash at conspecific receivers but which are not easily visible to raptors, which are a significant predator of lizards. Conversely, many species will balance the costs of sexual and natural selection and this is the case in Trinidadian guppies where populations living in predator-free streams are far more conspicuous than in predator-dense streams. These examples are from animals with static colour signals. A parallel solution that reduces the risk of exposure to predators is dynamic colour change because signals can be switched on and off relatively rapidly through neural or endocrine control. A wide range of invertebrates (e.g. cephalopods) and vertebrates (e.g. lizards and frogs) have convergently evolved this ability.

Unlike animals with static colour signals, we have a poor understanding of how animals with dynamic colour change respond to predation pressure and whether relaxing predation pressure will result in an increasingly conspicuous signal. Tackling this question in the wild is particularly difficult unless we can take advantage of unplanned changes to an organism’s environment such as during colonization of a novel habitat. Living organisms are occasionally translocated great distances by human action to environments far removed from their native range. These novel environments typically have a completely different suite of organisms, including predators, and offer a unique opportunity to study signal evolution in founder populations. Jackson’s chameleons (*Chamaeleo j. xanthopholus*) are just such a case. In 1972, a shipment of chameleons was sent from Kenya to the Hawaiian island of Oahu, destined for the pet trade. The animals arrived in poor condition and were left outdoors to recover, at which point they dispersed and became established on the island. On Oahu, the number of potential predators of chameleons is few (see methods), while in Kenya they are preyed upon by a wide range bird, snake and occasionally, mammal, predators.  This accidental introduction represents a unique opportunity to study the evolution of a dynamic visual signal by comparing native (Kenya) and introduced (Oahu) chameleons.

# Results

We manipulated both the social and predatory context chameleons encountered, in both their native Kenyan range and their introduced Hawaiian range, to test three hypotheses about colour signal evolution. Using spectrophotometry and visual modelling that relied on our understanding of the known chameleon visual system we quntified chromatic and achromatic contrast against the backgrounds in each environment (i.e., leaves and stems in Kenya and Hawaii).

Sexual selection is expected to favour more conspicuous dymamic sexual signals in Hawaii under relaxed predation because higher predation should favour cryptic signals whereas greater conspicuousness of sexual signals should be selected in social contexts. In support of this prediction, male display colouration exhibited in social contexts (i.e., male-male contests and courtship) was more conspicuous to a conspecific in Hawaii compared with Kenya (Male-Male Contests: Acromatic contrast ($dl$); $F_{`r Population_ConspicDisplayAC$df_num`,`r Population_ConspicDisplayAC$df_denom`}$ = `r Population_ConspicDisplayAC$Fstat`, p = `r p_value(Population_ConspicDisplayAC$p_val)`; Chromatic contrast ($dS$): $F_{`r Population_ConspicDisplayCC$df_num`,`r Population_ConspicDisplayCC$df_denom`}$ = `r Population_ConspicDisplayCC$Fstat`, p = `r p_value(Population_ConspicDisplayCC$p_val)`; Courtship displays: Acromatic contrast ($dl$); $F_{`r Population_ConspicCourtshipAC$df_num`,`r Population_ConspicCourtshipAC$df_denom`}$ = `r Population_ConspicCourtshipAC$Fstat`, p = `r p_value(Population_ConspicCourtshipAC$p_val)`; Chromatic contrast ($dS$): $F_{`r Population_ConspicCourtshipCC$df_num`,`r Population_ConspicCourtshipCC$df_denom`}$ = `r Population_ConspicCourtshipCC$Fstat`, p = `r p_value(Population_ConspicCourtshipCC$p_val)`), and this was  true across all body regions measured (See Supplemental Materials). Kenyan chameleons from populations where predation is high were significantly more crypic compared to introduced Hawaiian counterparts (Acromatic Contrast between Hawaiian and Kenyan Populations, Male-male contests: $\beta$ = `r robust_ConspicDisplayAC$beta[3]`, 95% CI = `r robust_ConspicDisplayAC_CI$CI_L[3]` to  `r robust_ConspicDisplayAC_CI$CI_U[3]`, df = `r robust_ConspicDisplayAC$df[3]`, t =  `r robust_ConspicDisplayAC$tstat[3]`, p = `r p_value(robust_ConspicDisplayAC$p_Satt[3])`, Courtship: $\beta$ = `r robust_ConspicCourtshipAC_CI$beta[3]`, 95% CI = `r robust_ConspicCourtshipAC_CI$CI_L[3]` to  `r robust_ConspicCourtshipAC_CI$CI_U[3]`, df = `r robust_ConspicCourtshipAC_CI$df[3]`, t =  `r summary(ConspicCourtshipAC)$coefficients[3, "t value"]`, p = `r p_value(robust_ConspicCourtshipAC$p_Satt[3])`, see figure 1a & b).

As a consequence of signal release, sexual signals are expected to be more easily detected by predators (i.e., snakes and birds) in Hawaii. Differential predation pressure resulting from snake and bird predation varying would also result in populations varying in their conspicuousness to different predators.  In support of these predictions, Hawaiian chameleons were significantly more conspicous to both snake and bird predators (figure 1b & c; Achromatic Contrast: ($dl$); $F_{`r Population_AC_Pred$df_num`,`r Population_AC_Pred$df_denom`}$ = `r Population_AC_Pred$Fstat`, p = `r p_value(Population_AC_Pred$p_val)`). Differences in chromatic contrast between populations, however,  depended on the predator (significant predator by population interaction: $F_{`r back_pred_CC_Pred$df_num`,`r back_pred_CC_Pred$df_denom`}$ = `r back_pred_CC_Pred$Fstat`, p = `r p_value(back_pred_CC_Pred$p_val)`). Interestingly, Kenyan and Hawaiian lizards were equally detectable by snakes, but Kenyan chameleons were significantly more cryptic from birds compared to Hawaiian chamleons (figure 1c; Snakes (Hawaii-Kenya): `r CI_hyp_test[1,2]`, 95% CI = `r CI_hyp_test[1,3]` to `r CI_hyp_test[1,4]`, p = 1; Birds (Hawaii-Kenya): `r CI_hyp_test[2,2]`, 95% CI = `r CI_hyp_test[2,3]` to `r CI_hyp_test[2,4]`, p = `r CI_hyp_test[2,8]`).

Local adaptation of sexual signals would result in Hawaiian chameleons being more conspicous to conspecifics in the Hawaiian environment compared to their native Kenyan environment. While it was not possible to conduct common garden experiments to provide definitive tests of local adaptation, our results support the local adaptation hypothesis given that Hawaiian chameleons were more conspicous to other chameleons in the Hawaiian environment compared to the same chameleons in a Kenyan environment in both social contexts (figure 2; Contrast between Hawaiian lizard on Kenyan vs Hawaiian on Hawaiin background - Male-male contests: Achromatic ($dl$), $\beta$ = `r CI_dis_AC[1,1]`, 95% CI = `r CI_dis_AC[1,2]` to `r CI_dis_AC[1,3]`, p = `r p_displ_AC`, Chromatic ($dS$) $\beta$ = `r CI_Disp_CC[1,1]`, 95% CI = `r CI_Disp_CC[1,2]` to `r CI_Disp_CC[1,3]`, p = `r p_displ_CC`; Courtship: Achromatic ($dl$), $\beta$ = `r CI_AC_Courtship[1,1]`, 95% CI = `r CI_AC_Courtship[1,2]` to `r CI_AC_Courtship[1,3]`, p = `r p_courtship_AC`, Chromatic ($dS$) $\beta$ = `r CI_CC_Courtship[1,1]`, 95% CI = `r CI_CC_Courtship[1,2]` to `r CI_CC_Courtship[1,3]`, p = `r p_courtship_CC`).

```{r Figure1, fig.width = 14, fig.height= 14, fig.cap= "Chromatic and achromatic contrast in Just Noticiable Differences (JND), of Hawaiian and Kenyan chameleons against their respective backgrounds (i.e., average environment of stems and leaves). A & B) Contrast of male chameleons during male-male contests and courtship. JNDs are calculated based on the chameleon visual system. NS denotes non-significant contasts, whereas astericks indicates significant differences is < 0.001. Insets showing spectral reflectance curves of two representative body regions ('gular' and 'top-flank), along with average signalling background (averaged leaves) are provided for each social and predatory context. A) spectral curves of signals during male-male displays; B) spectral curves of signals during courtship displays; C) spectral curves of signals during bird encounters; and D) spectral curves of signals during snake encounters."}
rerun = FALSE
if(rerun){
## Make figures that average across body regions, but display Social Context

@@

```{r Figure2, fig.width = 14, fig.height= 9, fig.cap= "Local adaptation of chromatic and acromatic contrast for Hawaiian chameleons against Hawaiian and Kenyan backgrounds."}
rerun=FALSE

>>>>

---
title: "Rapid evolution of a conspicuous dynamic visual signal in an introduced chameleon"
author: Whiting et al. 
date: "`r Sys.Date()`"
bibliography: refs.bib
csl: science.csl
output: 
  bookdown::word_document2:
    toc: no
    toc_depth: 6
    number_sections: false
    reference_docx: template.docx
  bookdown::html_document2:
    code_folding: hide
    number_sections: no
    toc: yes
    toc_depth: 6
    toc_float: yes
    bibliography: refs.bib
editor_options: 
  chunk_output_type: console
---

@@

```{r getlibrary, echo = FALSE, results = "hide"}
# load libraries
devtools::install_github("karthik/wesanderson")
pacman::p_load(lme4, lmerTest, ggplot2, lsmeans, magick, emmeans, multcomp, dplyr, wesanderson, patchwork, latex2exp, png, cowplot, magick, jpeg, clubSandwich, flextable)
source("./R/func.R")

@@

    # Check robustness of results to non-independence using Robust Variance Estimator (RVE)
    Display_1_IDs <- Display_1 %>% filter(!is.na(sc_SVL), !is.na(Population)) # filter out missing data
       robust_ConspicDisplayAC <- coef_test(ConspicDisplayAC, 
                                            cluster = Display_1_IDs$ID, vcov = "CR2")  
    robust_ConspicDisplayAC_CI <- conf_int(ConspicDisplayAC, 
                                           cluster = Display_1_IDs$ID, vcov = "CR2")

@@

 # Check robustness of results to non-independence using Robust Variance Estimator (RVE)
    Courtship_IDs <- Courtship %>% filter(!is.na(sc_SVL), !is.na(Population)) # filter out missing data
    robust_ConspicCourtshipAC <- coef_test(ConspicCourtshipAC, 
                                           cluster = Courtship_IDs$ID, vcov = "CR2")
 robust_ConspicCourtshipAC_CI <- conf_int(ConspicCourtshipAC, 
                                          cluster = Courtship_IDs$ID, vcov = "CR2")

@@

  # Check robustness of results to non-independence using Robust Variance Estimator (RVE)
  robust_localAdapt_DispBgrdAC <- clubSandwich::conf_int(DispBgrdAC, cluster = DisplayBgrd$ID, vcov =  "CR2")

@@

Martin J. Whiting <sup>1,*<sup>, Brenden S. Holland <sup>2<sup>, J. Scott Keogh <sup>3<sup>, Daniel W.A.Noble <sup>3<sup>, Katrina J. Rankin <sup>4<sup> and Devi M. Stuart-Fox <sup>4<sup>

1 Department of Biological Sciences, Macquarie University, Sydney, NSW 2109, Australia
2 Department of Natural Sciences, Hawaii Pacific University, Honolulu, Hawaii, USA
3 Division of Ecology and Evolution, Research School of Biology, The Australian National University, Canberra, ACT 2602, Australia
4 Department of Zoology, The University of Melbourne, Melbourne, Victoria 3010, Australia

# Abstract
Rapid evolution frequently follows a significant change to an organism’s environment, such as during colonization of a novel habitat. We show that Jackson’s chameleons (*Trioceros j. xanthopholus*) inadvertently introduced from Kenya to Hawaii (Oahu) have evolved significantly more conspicuous sexual signals. Compared to Kenya, Oahu has few potential predators and therefore, very low predation pressure. We used visual modelling to quantify the conspicuousness of wild chameleons from both Hawaii and Kenya to multiple social and predator stimuli in order to test for character release of sexual signals. The strongest effect was for achromatic contrast, where Hawaiian chameleons were brighter during male contests and courtship and were more conspicuous to snake and bird predators. Chameleon signals also showed evidence of local adaptation. Relaxing natural selection can result in signal release in systems experiencing strong sexual selection, including animals with dynamic colour change, and may act differently on the two primary visual channels (chromatic and achromatic).

@@

When Darwin tackled the joint problems of speciation and evolutionary change through natural selection, he was of the view that evolution was a gradual process that took place over extended time periods (1). We now have numerous examples of evolutionary change following natural selection in both the laboratory and wild, and not only in invertebrates with short generation times. Vertebrates, including Caribbean *Anolis* lizards and Galapagos finches, have demonstrated a remarkable capacity for morphological change of functional traits (limb length, beak size) on ecological time scales (e.g. (2, 3)). Furthermore, a far more recent development since Darwin’s seminal work on natural selection is the recognition that sexual selection can also drive significant changes in behaviour and phenotype over short time intervals (e.g., (4-6)).

Conspicuous colour signals evolve because of the fitness advantage conferred to the signaller. These signals frequently become increasingly elaborate through a combination of male contest competition and female preference (e.g. runaway selection (7, 8)). However, trait exaggeration is typically halted by natural selection because there is a threshold beyond which conspicuousness is too costly (9). Evolution can respond to this selection pressure in a variety of ways that include shifting conspicuous signals to areas of the body that might be concealed to predators. For example, many conspicuous lizards have colourful throats and venters that they flash at conspecific receivers but which are not easily visible to raptors, which are a significant predator of lizards (10). Conversely, many species will balance the costs of sexual and natural selection and this is the case in Trinidadian guppies where populations living in predator-free streams are far more conspicuous than in predator-rich streams (6, 9). These examples are from animals with static colour signals. A parallel solution that reduces the risk of exposure to predators is dynamic colour change because signals can be switched on and off relatively rapidly through neural or endocrine control (11, 12). A wide range of invertebrates (e.g. cephalopods) and vertebrates (e.g. lizards and frogs) have convergently evolved this ability (13-15).

Unlike animals with static colour signals, we have a poor understanding of how animals with dynamic colour change respond to predation pressure and whether relaxing predation pressure will result in an increasingly conspicuous signal. Tackling this question in the wild is particularly difficult unless we can take advantage of unplanned changes to an organism’s environment such as during colonization of a novel habitat. Living organisms are occasionally translocated great distances by human action to environments far removed from their native range. These novel environments typically have a completely different suite of organisms, including predators, and offer a unique opportunity to study signal evolution in founder populations. Jackson’s chameleons (*Trioceros j. xanthopholus*) are just such a case (Fig.\@ref(fig:figure1) and fig. S1). In 1972, a shipment of chameleons was sent from Kenya to the Hawaiian island of Oahu, destined for the pet trade. The animals arrived in poor condition and were left outdoors to recover, at which point they dispersed and became established on the island (16). On Oahu, the number of potential predators of chameleons is very few (no snakes or lizard-eating raptors), while in Kenya they are preyed upon by a wide range bird, snake and occasionally, mammal, predators. This accidental introduction represents a unique opportunity to study the evolution of a dynamic visual signal by comparing native (Kenya) and introduced (Oahu) chameleons under low and high levels of natural selection. Specifically, we tested the hypothesis that relaxed predation pressure results in character release of a dynamic visual signal. We tested the following predictions that underpin this hypothesis: 1.) display colouration used in male-male contests and courtship will be more conspicuous to a conspecific in Hawaii than Kenya; and 2.) display colouration used in response to an encounter with a bird and snake predator will be more conspicuous to both predator types in Hawaii than Kenya. We also asked whether colour signals were consistent with local adaptation. Specifically, we predicted that chameleon display colouration in each population would be more conspicuous within its own signaling environment compared to the other population’s signaling environment. For example, Hawaiian chameleons were historically found in Kenya but under local adaptation they are predicted to be more conspicuous in their current signaling environment, Hawaii. Likewise, Kenyan chameleons have no history in Hawaii and are therefore predicted to be more conspicuous within their current signaling environment, Kenya (table S1).

To test for character release and whether color signals show local adaptation, we quantified the local signalling environment (fig. S2) and manipulated the social and predatory context chameleons encountered, in both their native Kenyan, and introduced Hawaiian, range (fig. S3). We presented wild-caught male chameleons with the following treatments: a male, a female, a model bird predator (African cuckoo-hawk, *Aviceda cuculoides*) and a model snake predator (boomslang, *Dispholidus typus*) (Fig.\@ref(fig:figure1) and fig. S1,S4). We measured the spectral reflectance of males in dominant, subordinate, courtship, and antipredator (bird, snake) display states (Fig.\@ref(fig:figure2) and fig. S5). We then used visual modelling against a known chameleon visual system for sexual signals, and a bird and snake visual system for predation trials, to quantify chromatic and achromatic contrast against the backgrounds of each environment (supplementary text).

Under relaxed predation, sexual selection is expected to favour more conspicuous dynamic sexual signals in Hawaii than Kenya.  In support of this prediction, male display colouration exhibited in social contexts (i.e., male-male contests and courtship) was significantly more conspicuous with respect to brightness (Fig XX) to a conspecific in Hawaii compared with Kenya (male-male contests: achromatic contrast ($dl$); $F_{`r Population_ConspicDisplayAC$df_num`,`r Population_ConspicDisplayAC$df_denom`}$ = `r Population_ConspicDisplayAC$Fstat`, p = `r p_value(Population_ConspicDisplayAC$p_val)`; courtship displays: achromatic contrast ($dl$); $F_{`r Population_ConspicCourtshipAC$df_num`,`r Population_ConspicCourtshipAC$df_denom`}$ = `r Population_ConspicCourtshipAC$Fstat`, p = `r p_value(Population_ConspicCourtshipAC$p_val)`), but not significantly different in chromatic contrast ( Fig XX, male-male contests: $F_{`r Population_ConspicDisplayCC$df_num`,`r Population_ConspicDisplayCC$df_denom`}$ = `r Population_ConspicDisplayCC$Fstat`, p = `r p_value(Population_ConspicDisplayCC$p_val)`; courtship: $F_{`r Population_ConspicCourtshipCC$df_num`,`r Population_ConspicCourtshipCC$df_denom`}$ = `r Population_ConspicCourtshipCC$Fstat`, p = `r p_value(Population_ConspicCourtshipCC$p_val)`), and this was true across all body regions measured (See supplementary analyses). Kenyan chameleons, from a population where predation pressure is much higher, were significantly more crypic compared to their introduced Hawaiian counterparts (acromatic contrast between Hawaiian and Kenyan Populations, male-male contests: $\beta$ = `r robust_ConspicDisplayAC$beta[3]`, 95% CI = `r robust_ConspicDisplayAC_CI$CI_L[3]` to  `r robust_ConspicDisplayAC_CI$CI_U[3]`, df = `r robust_ConspicDisplayAC$df[3]`, t =  `r robust_ConspicDisplayAC$tstat[3]`, p = `r p_value(robust_ConspicDisplayAC$p_Satt[3])`, courtship: $\beta$ = `r robust_ConspicCourtshipAC_CI$beta[3]`, 95% CI = `r robust_ConspicCourtshipAC_CI$CI_L[3]` to  `r robust_ConspicCourtshipAC_CI$CI_U[3]`, df = `r robust_ConspicCourtshipAC_CI$df[3]`, t =  `r summary(ConspicCourtshipAC)$coefficients[3, "t value"]`, p = `r p_value(robust_ConspicCourtshipAC$p_Satt[3])`, see Figs XXXX).

As a consequence of signal release, sexual signals are expected to be more easily detected by predators (i.e., snakes and birds) in Hawaii. Also, differential predation pressure resulting from differences in predation rates by snake and bird predators would also result in populations varying in their conspicuousness to each predator type.  In support of these predictions, Hawaiian chameleons were significantly more conspicous to both snake and bird predators with respect to brightness (FigXX; achromatic contrast: ($dl$); $F_{`r Population_AC_Pred$df_num`,`r Population_AC_Pred$df_denom`}$ = `r Population_AC_Pred$Fstat`, p = `r p_value(Population_AC_Pred$p_val)`). Differences in chromatic contrast between populations, however, depended on the predator (significant predator by population interaction: $F_{`r back_pred_CC_Pred$df_num`,`r back_pred_CC_Pred$df_denom`}$ = `r back_pred_CC_Pred$Fstat`, p = `r p_value(back_pred_CC_Pred$p_val)`). Interestingly, Kenyan and Hawaiian chameleons were equally detectable by snakes, but Hawaiian chameleons were significantly more conspicuous to birds compared to Kenyan chameleons (FIG XX; Snakes (Hawaii-Kenya): `r CI_hyp_test[1,2]`, 95% CI = `r CI_hyp_test[1,3]` to `r CI_hyp_test[1,4]`, p = 1; Birds (Hawaii-Kenya): `r CI_hyp_test[2,2]`, 95% CI = `r CI_hyp_test[2,3]` to `r CI_hyp_test[2,4]`, p = `r CI_hyp_test[2,8]`).

Local adaptation of sexual signals would result in Hawaiian chameleons being more conspicous to conspecifics in the Hawaiian environment compared to their native Kenyan environment. While it was not possible to conduct common garden experiments to provide definitive tests of local adaptation, our results support the local adaptation hypothesis. Hawaiian chameleons were more conspicuous to other chameleons against heir own (Hawaiian) background as opposiued to the Kenyan background, in both social contexts and for both achromatic and chromatic contrast (FIG XX; Contrast between Hawaiian lizard on Kenyan vs Hawaiian on Hawaiin background - male-male contests: achromatic ($dl$), $\beta$ = `r CI_dis_AC[1,1]`, 95% CI = `r CI_dis_AC[1,2]` to `r CI_dis_AC[1,3]`, p = `r p_displ_AC`, chromatic ($dS$) $\beta$ = `r CI_Disp_CC[1,1]`, 95% CI = `r CI_Disp_CC[1,2]` to `r CI_Disp_CC[1,3]`, p = `r p_displ_CC`; courtship: achromatic ($dl$), $\beta$ = `r CI_AC_Courtship[1,1]`, 95% CI = `r CI_AC_Courtship[1,2]` to `r CI_AC_Courtship[1,3]`, p = `r p_courtship_AC`, chromatic ($dS$) $\beta$ = `r CI_CC_Courtship[1,1]`, 95% CI = `r CI_CC_Courtship[1,2]` to `r CI_CC_Courtship[1,3]`, p = `r p_courtship_CC`).

We have documented a surprisingly strong effect of environment on the evolution of a conspicuous, dynamic sexual signal. For a founding population of chameleons, the tropical environment of Hawaii represents a particularly favourable set of conditions with respect to habitat, temperature, and rainfall. Furthermore, Jackson’s chameleons are dietary generalists (17) and likely experience low competition for relatively abundant arthropod prey. But most importantly, Hawaii has no snakes and no raptors that feed on lizards. Domestic and feral cats would be their main threat on Oahu, and cats have very limited colour vision (18). In the absence of the usual suite of predators they would experience in their native rage of Kenya—a diverse community of snake and bird predators, the conditions were highly favourable for character release of sexual signals—particularly in a polygynous species in which males have intense male contest competition and elaborate courtship displays (Fig.\@ref(fig:figure1)). Not only do males use colorful displays, but they have three horns that they lock and twist into rivals during intense bouts that sometimes result in physical injury. Consequently, a ‘perfect storm’ of conditions set the stage for what we have demonstrated—an increasingly bright sexual signal used in both male contests and during courtship.

Evolutionary biologists, particularly those interested in speciation, seldom consider the role of natural selection in shaping sexually selected signals even though sexual signals evolve in an ecological context (19) (but see (5, 20, 21). Landmark studies such as those on Trinidadian guppies (6, 9) have shown the important interaction between natural and sexual selection in static signals following experimental manipulation. Ours is one of the only studies that show that a dynamic sexual signal—rapid colour change—is likewise constrained by natural selection. Our results suggest that dynamic sexual signals are not simply plastic in a way that allows an individual to moderate the extent to which they ‘turn on’ a display but rather, that there is a threshold constrained by natural selection. This may in part be explained by the nature of colour change in chameleons. Chameleons have two layers of superposed iridiophores that facilitates color change that produces both camouflage and conspicuous sexual signals (22). Interestingly, while Hawaiian chameleons have evolved brighter display coloration, other elements of colour including hue and saturation have not experienced similar levels of selection except to become more conspicuous against their own background (local adaptation). This suggests that the two visual channels (achromatic vs chromatic) experience differing levels of selection which is perhaps not surprising given that they act independently of one another (23). While studying evolution is challenging because it is not possible to go back in time to witness key evolutionary events, we can take advantage of unique opportunities such as an invasion, to better understand how natural and sexual selection interact to shape animal signals.

```{r figure1, fig.cap = "Chameleon colour signal change in response to different social and predatory stimuli. Male chameleons experience intense sexual selection. During the breeding season they change from a neutral state (green) to a highly conspicuous bright yellow display signal. They also readily fight by locking horns and sometimes pierce their rivals skin with their horns. (A) A typical dominant male in display coloration. (B) A subordinate male that lost a contest and turned from bright yellow to brown. (C) Two males fighting, both are in display colouration and at a stage where they are relatively even matched. (D) A male courting female in full display color while the female has turned to a contrasting colour, which suggests rejection of the male"}

img1 <- readJPEG("./photos for figs/MS_images/Fig1a.jpg")
img2 <- readJPEG("./photos for figs/MS_images/Fig1b.jpg")
img3 <- readJPEG("./photos for figs/MS_images/Fig1c.jpg")
img4 <- readJPEG("./photos for figs/MS_images/Fig1d.jpg")

p1 <- cowplot::ggdraw()+
        cowplot::draw_image(img1)
p2 <- cowplot::ggdraw()+
        cowplot::draw_image(img2)
p3 <- cowplot::ggdraw()+
        cowplot::draw_image(img3)
p4 <- cowplot::ggdraw()+
        cowplot::draw_image(img4)

(p1 + p2) / (p3 + p4) + plot_annotation(tag_levels = c("A", "B", "C", "D"))

```{r figure2, fig.width = 14, fig.height= 14, fig.cap="Mean spectral reflectance curves for male chameleons for representative body regions (gular, top flank) and background (leaves) for Hawaii and Kenya. The context for measurement was (A) male contest displays; (B) courtship; (C) bird predator; and (D) snake predator. More details in online supplementary material and Fig. S5"}
png <- image_read("./output/figures/Figure3.png")
png

```{r figure3, fig.width = 14, fig.height= 14, fig.cap= "Chromatic and achromatic contrast in Just Noticiable Differences (JND), of Hawaiian and Kenyan chameleons against their respective backgrounds (i.e., average environment of stems and leaves). (A & B) Contrast of male chameleons during male-male contests and courtship. JNDs are calculated based on the chameleon visual system. NS denotes non-significant contrasts, whereas asterisks indicates significant differences where p is less than 0.001. Insets showing spectral reflectance curves of two representative body regions ('gular' and 'top-flank'), along with average signalling background (averaged leaves) are provided for each social and predatory context. (A) spectral curves of signals during male-male displays; (B) spectral curves of signals during courtship displays; (C) spectral curves of signals during bird encounters; and (D) spectral curves of signals during snake encounters."}
rerun = FALSE
if(rerun){
## Make figures that average across body regions, but display Social Context

@@

```{r figure4, fig.width = 14, fig.height= 9, fig.cap= "Evidence for local adaptation of color signals in both (A) chromatic- and (B) achromatic contrast for Hawaiian chameleons. For local adaptation, signals are predicted to be more conspicuous against their own background. In this case, Hawaiian chameleons were both significantly more chromatic and brighter (achromatic contrast)  against their own background compared to the Kenyan background suggesting that selection has acted on increasing local signal conspicuousness."}
rerun=FALSE
